ECCCos from the Black Box:\\ Faithful Explanations through\\ Energy-Constrained Conformal Counterfactuals

### 2023-09-20 paper/appendix.tex
slowly slowly
just missing section 6
more shit in the app

Consistent with~\citet{grathwohl2020your}, we have specified $\epsilon=2$ and $\sigma=0.01$ as the default values for all of our experiments. The number of total SGLD steps $J$ varies by dataset (Table~\ref{tab:ebmparams}). Following best practices, we initialize $\mathbf{x}_0$ randomly in 5\% of all cases and sample from a buffer in all other cases. The buffer itself is randomly initialised and gradually grows to a maximum of 10,000 samples during training as $\hat{\mathbf{x}}_{J}$ is stored in each epoch~\citep{du2020implicit,grathwohl2020your}.

@@

\begin{equation}\label{eq:jem-loss}
  \begin{aligned}
    L(\theta) &= L_{\text{clf}}(\theta) + L_{\text{gen}}(\theta) + \lambda L_{\text{reg}}(\theta) 
  \end{aligned}
\end{equation}

@@

\subsubsection{Differentiability}

@@

\subsection{ECCCo}\label{app:eccco}

In this section, we briefly discuss convergence conditions for CE and provide details concerning the actual implementation of our framework in \texttt{Julia}.  
\subsubsection{A Note on Convergence}

@@

Table~\ref{tab:params} provides an overview of all parameters related to our experiments. The \textit{GMSC} data were randomly undersampled for balancing purposes and all features were standardized. \textit{MNIST} data was also randomly undersampled for reasons outlined below. Pixel values were preprocessed to fall in the range of $[-1,1]$ and a small Gaussian noise component ($\sigma=0.03$) was added to training samples following common practice in the EBM literature. All of our models were trained through mini-batch training using the Adam optimiser (\citet{kingma2017adam}). Table~\ref{tab:perf} shows standard evaluation metrics measuring the predictive performance of our different models grouped by dataset. These measures were computed on test data.

>>>>

Consistent with~\citet{grathwohl2020your}, we have specified $\epsilon=2$ and $\sigma=0.01$ as the default values for all of our experiments. The number of total SGLD steps $J$ varies by dataset (Table~\ref{tab:ebmparams}). Following best practices, we initialize $\mathbf{x}_0$ randomly in 5\% of all cases and sample from a buffer in all other cases. The buffer itself is randomly initialised and gradually grows to a maximum of 10,000 samples during training as $\hat{\mathbf{x}}_{J}$ is stored in each epoch~\citep{du2019implicit,grathwohl2020your}.

@@

\begin{equation}\label{eq:jem-loss}
  \begin{aligned}
    L_{\text{JEM}}(\theta) &= L_{\text{clf}}(\theta) + L_{\text{gen}}(\theta) + \lambda L_{\text{reg}}(\theta) 
  \end{aligned}
\end{equation}

@@

\subsubsection{Differentiability}\label{app:cp-diff}

@@

\subsection{\textit{ECCCo}}\label{app:eccco}

In this section, we explain \textit{ECCCo} in some more detail, briefly discuss convergence conditions for counterfactual explanations and provide details concerning the actual implementation of our framework in \texttt{Julia}.

\subsubsection{More detail on our generator}

The counterfactual search objective for \textit{ECCCo} was introduced in Equation~\ref{eq:eccco} in the body of the paper. We restate this equation here for reference:

\begin{equation} \label{eq:eccco-app}
  \begin{aligned}
  \mathbf{Z}^\prime &= \arg \min_{\mathbf{Z}^\prime \in \mathcal{Z}^L} \{  {\text{yloss}(M_{\theta}(f(\mathbf{Z}^\prime)),\mathbf{y}^+)}+ \lambda_{1} {\text{dist}(f(\mathbf{Z}^\prime),\mathbf{x}) } \\
  &+ \lambda_2 \Delta\mathcal{E}(\mathbf{Z}^\prime,\widehat{\mathbf{X}}_{\theta,\mathbf{y}^+}) + \lambda_3 \Omega(C_{\theta}(f(\mathbf{Z}^\prime);\alpha)) \} 
  \end{aligned} 
\end{equation}

We can make the connection to energy-based modeling more explicit by restating this equation in terms $L_{\text{JEM}}(\theta)$, which we defined in Equation~\ref{eq:jem-loss}. In particular, note that for $\lambda_2=1$ and $\lambda L_{\text{reg}}(\theta)=0$ we have

\begin{equation} \label{eq:eccco-jem}
  \begin{aligned}
  \mathbf{Z}^\prime &= \arg \min_{\mathbf{Z}^\prime \in \mathcal{Z}^L} \{  {L_{\text{JEM}}(\theta;M_{\theta}(f(\mathbf{Z}^\prime)),\mathbf{y}^+)}+ \lambda_{1} {\text{dist}(f(\mathbf{Z}^\prime),\mathbf{x}) }  + \lambda_3 \Omega(C_{\theta}(f(\mathbf{Z}^\prime);\alpha)) \} 
  \end{aligned} 
\end{equation}

since $\Delta\mathcal{E}(\cdot)$ is equivalent to the generative loss function $L_{\text{gen}}(\cdot)$. In fact, this is also true for $\lambda L_{\text{reg}}(\theta)\ne0$ since we use the Ridge penalty $L_{\text{reg}}(\theta)$ in the counterfactual search just like we do in joint-energy training. This detail was omitted from the body of the paper for the sake of simplicity.

Aside from the additional penalties in Equation~\ref{eq:eccco-app}, the only key difference between our counterfactual search objective and the joint-energy training objective is the parameter that is being optimized. In joint-energy training we optimize the objective with respect to the network weights $\theta$:

\begin{equation}\label{eq:jem-grad}
  \begin{aligned}
    \nabla_{\theta} L_{\text{JEM}}(\theta) &=\nabla_{\theta} L_{\text{clf}}(\theta) + \nabla_{\theta}L_{\text{gen}}(\theta) + \lambda \nabla_{\theta} L_{\text{reg}}(\theta) 
  \end{aligned}
\end{equation}

we take these parameters as fixed in our counterfactual search.

\subsubsection{A Note on Convergence}

@@

Table~\ref{tab:params} provides an overview of all parameters related to our experiments. The \textit{GMSC} data were randomly undersampled for balancing purposes and all features were standardized. \textit{MNIST} data was also randomly undersampled for reasons outlined below. Pixel values were preprocessed to fall in the range of $[-1,1]$ and a small Gaussian noise component ($\sigma=0.03$) was added to training samples following common practice in the EBM literature. All of our models were trained through mini-batch training using the Adam optimiser (\citet{kingma2014adam}). Table~\ref{tab:perf} shows standard evaluation metrics measuring the predictive performance of our different models grouped by dataset. These measures were computed on test data.
